A generic and fast C++ optimization framework

### 2017-10-22 paper.tex
Add affiliation and email address
Capitalize library names. Change expressions.

\author{
  Shikhar Bhardwaj \\
  \AND
  Yannis Mentekidis \\
  \AND
  Marcus Edel \\
  Free University of Berlin \\
  Arnimallee 7, 14195 Berlin \\
  \texttt{marcus.edel@fu-berlin.de} \\
  \AND
  Ryan R. Curtin \\
  Symantec Corporation \\
  Atlanta, GA, USA 30318 \\
  \texttt{ryan@ratml.org} \\
}

@@

\begin{abstract}
The development of the {\bf mlpack} C++ machine learning library
(\url{http://www.mlpack.org/}) has required the design and implementation of a
flexible, robust optimization system that is able to solve the types of
arbitrary optimization problems that may arise all throughout machine learning
problems.  In this paper, we present the generic optimization framework that we
have designed for mlpack.  A key priority in the design was ease of
implementation of both new optimizers and new objective functions to be
optimized; therefore, implementation of a new optimizer requires only one
method and implementation of a new objective function requires at most four
functions.  This leads to simple and intuitive code, and for fast prototyping
and experimentation, this is of paramount importance.  When compared to
optimization frameworks of other libraries, we find that mlpack's supports more
types of objective functions, is able to make optimizations that other
frameworks do not, and seamlessly supports user-defined objective functions
{\it and} optimizers.
%\begin{itemize}
%  \item It's easy to implement new optimizers.
%  \item It's easy to implement new functions to be optimized.
%  \item This system is more comprehensive than other techniques that may focus
%only on neural networks or single problems.
%  \item This system is also more specific and offers more room for optimization
%than super general systems like scipy.optimize().
%\end{itemize}
%
%This abstract will need to be rewritten.
\end{abstract}

@@

Given this situation, we have developed a flexible optimization infrastructure
in the {\bf mlpack} C++ machine learning library \cite{mlpack2013}.  This
infrastructure makes it easy to combine nearly any type of optimizer with nearly
any type of objective function, and has allowed us to minimize the effort
necessary to both implement new optimizers and to implement new machine learning
algorithms that depend on optimization.  Since the framework is implemented in
C++ and uses template metaprogramming, we are able to preserve clean syntax
while simultaneously allowing compile-time optimizations that can result in a
generic optimizer that is as fast as an optimizer written for one specific objective function.

@@

        & shogun \cite{sonnenburg2010shogun}             & \CIRCLE & - & \CIRCLE & \CIRCLE & \CIRCLE & - & -

@@

        & vowpal wabbit \cite{Langford2007VW}      & - & - & \CIRCLE  & - & - & - &

@@

        & caffe \cite{jia2014caffe}           & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle

@@

        & keras \cite{chollet2015}            & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle

@@

        & scipy \cite{jones2014scipy}             & \CIRCLE & \CIRCLE  & -  & \CIRCLE & - & - & - \\

@@

Both of these methods should calculate the gradient and place the results into
the matrix object {\tt gradient}.  Note that the method accepts a template
parameter {\tt GradType}, which may be {\tt arma::mat} (dense Armadillo matrix)
or {\tt arma::sp\_mat} (sparse Armadillo matrix).  This allows support for both
sparse-supporting and non-sparse-supporting optimizers.\footnote{One could write
a non-templated {\tt Gradient()} method for just {\tt arma::mat}, and it would
work fine for non-sparse optimizers.  But to us it seems just as easy to
templatize it.}

>>>>

\author{
  Shikhar Bhardwaj \\
  Delhi Technological University \\
  Delhi - 110042 \\
  \texttt{shikhar_bt2k15@dtu.ac.in} \\
  \AND
  Yannis Mentekidis \\
  \AND
  Marcus Edel \\
  Free University of Berlin \\
  Arnimallee 7, 14195 Berlin \\
  \texttt{marcus.edel@fu-berlin.de} \\
  \AND
  Ryan R. Curtin \\
  Symantec Corporation \\
  Atlanta, GA, USA 30318 \\
  \texttt{ryan@ratml.org} \\
}

@@

\begin{abstract}
The development of the {\bf mlpack} C++ machine learning library
(\url{http://www.mlpack.org/}) has required the design and implementation of a
flexible, robust optimization system that is able to solve the types of
arbitrary optimization problems that may arise all throughout machine learning
problems.  In this paper, we present the generic optimization framework that we
have designed for mlpack.  A key priority in the design was ease of
implementation of both new optimizers and new objective functions to be
optimized; therefore, implementation of a new optimizer requires only one
method and implementation of a new objective function requires at most four
functions.  This leads to simple and intuitive code, which, for fast prototyping
and experimentation, is of paramount importance.  When compared to
optimization frameworks of other libraries, we find that mlpack's supports more
types of objective functions, is able to make optimizations that other
frameworks do not, and seamlessly supports user-defined objective functions
{\it and} optimizers.
%\begin{itemize}
%  \item It's easy to implement new optimizers.
%  \item It's easy to implement new functions to be optimized.
%  \item This system is more comprehensive than other techniques that may focus
%only on neural networks or single problems.
%  \item This system is also more specific and offers more room for optimization
%than super general systems like scipy.optimize().
%\end{itemize}
%
%This abstract will need to be rewritten.
\end{abstract}

@@

Given this situation, we have developed a flexible optimization infrastructure
in the {\bf mlpack} C++ machine learning library \cite{mlpack2013}.  This
infrastructure makes it easy to combine nearly any type of optimizer with nearly
any type of objective function, and has allowed us to minimize the effort
necessary to both implement new optimizers and to implement new machine learning
algorithms that depend on optimization.  Since the framework is implemented in
C++ and uses template metaprogramming, we are able to preserve clean syntax
while simultaneously allowing compile-time speedups that can result in a
generic optimizer that is as fast as an optimizer written for one specific
objective function.

@@

        & Shogun \cite{sonnenburg2010shogun}             & \CIRCLE & - & \CIRCLE & \CIRCLE & \CIRCLE & - & -

@@

        & Vowpal Wabbit \cite{Langford2007VW}      & - & - & \CIRCLE  & - & - & - &

@@

        & Caffe \cite{jia2014caffe}           & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle

@@

        & Keras \cite{chollet2015}            & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle

@@

        & SciPy \cite{jones2014scipy}             & \CIRCLE & \CIRCLE  & -  & \CIRCLE & - & - & - \\

@@

Both of these methods should calculate the gradient and place the results into
the matrix object {\tt gradient} that is passed as an argument.
Note that the method accepts a template
parameter {\tt GradType}, which may be {\tt arma::mat} (dense Armadillo matrix)
or {\tt arma::sp\_mat} (sparse Armadillo matrix).  This allows support for both
sparse-supporting and non-sparse-supporting optimizers.\footnote{One could write
a non-templated {\tt Gradient()} method for just {\tt arma::mat}, and it would
work fine for non-sparse optimizers.  But to us it seems just as easy to
templatize it.}
